\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2020-09-28 redaction/revision_JEGeo/referees/answers/Answer_referee1_Lise.tex

\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}  We thank the referee for this valuable suggestion. It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.

\item \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.} We agree with the referee that the comparison with the literature (Hummels (2007) in particular) was not straightforward for the reader in the submitted version. We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method. In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
    \begin{equation}
    f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta_{ikt}} \label{eq:Hummels}
    \end{equation}

     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ik}$ the price-elasticity of transport costs, which also corresponds to the share of additive costs in total transport costs (as we now show in the paper's appendix):

     \begin{equation}
     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)}-1+\frac{t_{is(k)}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
     \end{equation}

@@

     As underlined in the paper (both submitted and revised versions), Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method, as he estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, yielding to estimates of $\beta_{is(k)}$ that are invariant over time (but mode-specific).

     Two alternative strategies are yet available. Let us start with the referee's method. It starts from the functional form linking transport costs and the fas price as specified in the Hummels's equation (\ref{eq:Hummels}). Taken in log (on a mode/yearly basis), it is written as:
     $$\ln f_{ikd} = \beta_{is(k)}\ln \widetilde{p}_{ikd} +\text{Controls}_{ikd}+ \varepsilon_{ikd}$$

     From this, one can then recover the levels of additive / multiplicative transports costs. Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ for a given sector-country $i,s(k)$, one can indeed solve the following two-equation system:

\begin{eqnarray}
p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
\end{eqnarray}

@@

\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode). With 2 equations and 2 endogenous variables, the system might be solved.

     Alternatively, our method rather starts from the definition of transport costs as:
     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis). Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.

### 2021-01-22 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement mineurs rédaction
la biblio ne compile pas

\documentclass[a4paper,12pt]{article}

@@

Trade costs are central to international economic analysis. In particular, they are
known to be a major obstacle to international economic integration and international trade
flows. Defined as the costs associated with the exchange of goods across
national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially
decreased with technological advance in transportation, infrastructure development
and new communication technologies. However, several empirical evidence
suggest that transport costs remain large and deserve attention.\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \citet{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. \citet{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from \textbf{1989 to 2019}. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\citet{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \citet{Irrazabal_2015}, or \citet{martin2012}). Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs. Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1989-2019 and by transport mode (air or maritime transport).\smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e. changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \citet{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet. First, we consider a larger period of time, spanning from 1974 (as in \citet{hummels2007}) to 2019 (while he stops in 2004). Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019). For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \citet{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs. In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs. Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. \textbf{A FAIRE DANS LES REGRESSIONS, METHODE Hummels mais avec s=3, k=10. Peut-on le faire? For the argument to be convincing, need to do it on the same years with the same sector-product degrees.}

\smallskip
In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}). As shown by \citet{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the empirical findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
Using transaction-level trade data from Colombia, \citet{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs). One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \citet{Alessandria-et-al-AER-2010} or \citet{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs.
Yet, our paper complements their findings in two main respects.
First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
While \citet{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from \textbf{1989 to 2019}.
In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since \textbf{1989}.

\textbf{Big Picture}.
Our results have important implications for both international trade and international macroeconomics models. Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \citet{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs. On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models. In both fields yet, not much progress has been made in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}). The quantitative results we provide in the paper allow to make one step in filling that gap. \textbf{To develop}. \medskip

@@

Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \citet{hummels2007}. The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2019 (details on the database provided in Appendix \ref{app:data}). We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price. The fas price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in. It is also the seller's responsibility to clear the goods for export. The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer. Once the goods clear the railing of the vessel, the buyer assumes the risk. Note that this term is used exclusively for maritime and inland waterway transport. While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, \textbf{both items being conditional to the transport mode}; in other words, it is the price for one kg of the goods net of transportation costs, \textbf{by transport mode}. The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. \textbf{Cf point 2 refere 1 sur unit prices; avoir quantite plutot que poids, robustesse a faire sur sous-echantillon des produits qui ne passent que par un seul mode de transport).} Our dependant variable is finally computed as the ratio of the cif price divided by the fas price. By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
This is quite a standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.

Obviously, this dataset has limitations.
First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
It is thus silent about the other dimensions of international trade costs.
Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
Further, the evidence summarized in \citet{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

Exploiting this dataset also has three advantages.
First, it comes from a single, homogenous and trustworthy customs source.
This limits the measurement error bias.
Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level (with a concordance code to the SITC coding system) from \textbf{1989 to 2019}. For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level. %This will be crucial to compute transport costs, as we detail below.
In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g. \citet{Irrazabal_2015}.
Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \citet{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs.
In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip

@@

\paragraph{Estimation strategy} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped. In accordance with \citet{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.). In this case, the link between the import, export prices and shipping charges is given by:

@@

\paragraph{Obtaining the estimated equation} We follow \citet{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $(1+\tau_i)\times (1+\tau_k) -1 \simeq \tau_i + \tau_k$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
We also provide a robustness analysis to the separability assumption.
Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries.
See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}

@@

\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair. In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.

@@

\subsection{Comparing with \citet{hummels2007} \label{app:compare_Hummels}}

>>>>

\documentclass[a4paper,11pt]{article}

@@

Trade costs are central to international economic analysis. In particular, they are
known to be a major obstacle to international economic integration and international trade
flows. Defined as the costs associated with the exchange of goods across
national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially
decreased with technological advance in transportation, infrastructure development
and new communication technologies. However, several empirical evidence
suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from \textbf{1989 to 2019}. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\cite{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}). Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs. Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1989-2019 and by transport mode (air or maritime transport).\smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs comes mainly from the reduction of transport cost \textit{per se}, , by roughly \textbf{50\% in air shipping and 60\% in vessel shipping over 1974-2013}.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects (i.e. changes over time in the baskets of imported goods and/or the origin countries).} This contrasts with the conclusion obtained by \cite{hummels2007} yet also on US imports flows.\footnote{Both Hummel's and our's databases come from the US Bureau of Census's import flows database, with two differences yet. First, we consider a larger period of time, spanning from 1974 (as in \cite{hummels2007}) to 2019 (while he stops in 2004). Second, while Hummels considers products at the 5-digit classification level, we are able to exploit heterogeneity in the data at the much more disaggregated HS-10 level for the vast majority of the period (1989-2019). For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level.}\textbf{Parler de marge extensive - intensive} \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs. In contrast, we find that trade composition effects have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs. Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turn out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. \textbf{A FAIRE DANS LES REGRESSIONS, METHODE Hummels mais avec s=3, k=10. Peut-on le faire? For the argument to be convincing, need to do it on the same years with the same sector-product degrees.}

\smallskip
In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}). As shown by \cite{alchian}, the relative price of two varieties of some goods will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher-quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the empirical findings by \cite{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity.\footnote{Also, their estimates imply that doubling freight costs increases average free alongside ship (fas) export prices by 80\% to 141\%, consistent with high-quality goods being sold in markets with high freight costs.}
Recent empirical studies based on micro-level data also provide strong empirical support for the role of additive costs (i.e. cost per unit exported) in international trade costs.\footnote{Based on a firm-product-level database of French exporters, \cite{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.
Using transaction-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.} If in empirical support for additive trade costs, these papers yet do not allow for a quantitative measure of these cost (neither for the ad-valorem costs). One important contribution of the paper is to provide quantitative estimates of both type of transport costs.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \cite{Kropf-Saure-JIE-2016}, who estimate the size and shape of per-shipment costs based on Swiss data, and to \cite{Alessandria-et-al-AER-2010} or \cite{Hornok-et-al-JIE-2015, Hornok-et-al-RES-2015}, who point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

Closely related to our paper is the work by \cite{Irrazabal_2015}, which develops a structural framework for inferring relative additive trade costs from firm-level trade data.
Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (also expressed in Norwegian crowns).
Our results share in common with \cite{Irrazabal_2015} the important role of the additive component of international trade costs.
Yet, our paper complements their findings in two main respects.
First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs.
From this, we can rebuild the ratio in similar terms to them, thereby gaining in generality in this respect.
While \cite{Irrazabal_2015} find that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our results point to a value paid in additive transport costs of \textbf{2.8 USD for maritime shipping and 1.5 USD for air transport for 2004.}
This comparison of results suggests that between \textbf{a third and a half of the additive trade costs} are attributable to the transport sector (depending on the transport mode).
Second, we exploit exhaustive information about the imports flows of the US over a wide time-span from \textbf{1989 to 2019}.
In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time.
In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since \textbf{1989}.

\textbf{Big Picture}.
Our results have important implications for both international trade and international macroeconomics models. Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs.\footnote{ \cite{sorensen2014} extends \cite{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}} These results suggest that important welfare gains can be achieved by reducing additive trade costs. On the same toke, as additive costs distort relative prices, they are very likely to affect the international transmission of shocks in international business cycles models. In both fields yet, not much progress has been made in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citealp{Irrazabal_2015}). The quantitative results we provide in the paper allow to make one step in filling that gap. \textbf{To develop}. \medskip

@@

Our analysis of transportation costs consists of exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007}. The database we use to construct our measure of transport costs comes from the US annual ``Imports of Merchandise'' provided by the US Census Bureau, spanning 1974 to 2019 (details on the database provided in Appendix \ref{app:data}). We first use customs values, quantities and freight costs to recover free alongside ship (fas) and cost-insurance-freight (cif) prices, by good, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob (free on board) price rather than the fas price. The fas price means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in. It is also the seller's responsibility to clear the goods for export. The FOB price means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer. Once the goods clear the railing of the vessel, the buyer assumes the risk. Note that this term is used exclusively for maritime and inland waterway transport. While both terms are closely related, the US Foreign Trade Statistics reports fas prices.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight, \textbf{both items being conditional to the transport mode}; in other words, it is the price for one kg of the goods net of transportation costs, \textbf{by transport mode}. The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. \textbf{Cf point 2 refere 1 sur unit prices; avoir quantite plutot que poids, robustesse a faire sur sous-echantillon des produits qui ne passent que par un seul mode de transport).} Our dependant variable is finally computed as the ratio of the cif price divided by the fas price. By construction higher than 1, the variable provides a measure of transport costs as a proportion of the goods' price, an ad-valorem equivalent.
This is quite a standard and widespread strategy, as emphasized by \cite{anderson_wincoop_jel}.

Obviously, this dataset has limitations.
First, it restricts our analysis to the study of international \emph{transport}  costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs.
It is thus silent about the other dimensions of international trade costs.
Second, in terms of transport costs \textit{per se}, our measure omits the ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size, etc.).
In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that these direct transport costs do represent a sizable share of trade costs.
According to \cite{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs that amount to a  9\% tax equivalent.
Further, the evidence summarized in \cite{anderson_wincoop_jel} points to the persisting importance of direct transport costs, especially compared to other trade barriers.
They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \cite{Hummels_1999} mentions several papers which all indicate that transport costs pose a barrier similar in size to, or larger than tariffs.
In the same vein, \cite{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

Exploiting this dataset also has three advantages.
First, it comes from a single, homogenous and trustworthy customs source.
This limits the measurement error bias.
Based on customs declarations, the US Imports Database inventories all imports (both values and quantities), by country of origin, to the United States at the HS 10-digit level (with a concordance code to the SITC coding system) from \textbf{1989 to 2019}. For the period 1974-1988, we rely on Hummel's original database with a 5-digit product classification level. %This will be crucial to compute transport costs, as we detail below.
In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, land/ocean (or ``vessel'') and air.
We will make use of this to highlight potential differences in the dynamics of transport costs across transportation mode.
Second, this dataset provides both the import price and the export price for the same good (for a given origin country).
This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to e.g. \cite{Irrazabal_2015}.
Third, this dataset is available over a long time span, which enables us to analyze the trend patterns of international transport costs over the period \textbf{1989-2019}.\footnote{Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures are based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs.
In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``potential'' transport costs.} \smallskip

@@

\paragraph{Estimation strategy} Our purpose is to provide time-varying estimates of the size of ad-valorem transport costs and additive transport costs.
To do so, denote by $p$ the price per kg of goods paid by the importer (import, or cif price), $\widetilde{p}$ the (per kg) export (or fas) price and $f$ the shipping charge per kg shipped. In accordance with \cite{hummels2010}, assume that shipping charges may depend on the price $\widetilde{p}$, such that per kg shipping charges are written as $f = \widetilde{p}^{1-\beta} X$, where $X$ represents all other elements of transport costs (such as distance, port quality, etc.). In this case, the link between the import, export prices and shipping charges is given by:

@@

\paragraph{Obtaining the estimated equation} We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $(1+\tau_i)\times (1+\tau_k) -1 \simeq \tau_i + \tau_k$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
We also provide a robustness analysis to the separability assumption.
Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries.
See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}

@@

\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price).
Besides, related concerns may arise from productivity-sorting model in \cite{melitz} or the quality-sorting model in \cite{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair. In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.

@@

\subsection{Comparing with \cite{hummels2007} \label{app:compare_Hummels}}
